### 2020-09-16 paper/implementation.tex
many many fixes

A repository containing the discussed source code and tests is available under the link: \href{https://github.com/AdrianSiwiec/Perfect-Graphs}{https://github.com/AdrianSiwiec/Perfect-Graphs}
\todo{make repo public and add readme on howto run, some script creating and running default tests}

@@

First of all, there are many methods for generating all possible paths to choose from. We could simply enumerate all sequences of vertices without repetition and for each one of them check if it is a path, i.e., if all pairs of vertices next to each other are connected and that there are no other edges, but this would be too slow and render our algorithm unusable.

@@

But we can do much better with some care and a better data structure. For a given graph we create a data structure that suits best the goal of generating paths. In addition to having for each vertex a list of its neighbors we create data structure that will allow us to generate a candidate for next path in amortized $O(1)$ time.

@@

In our use case, a few optimizations can be made. First, we don't need a hashing set. We can have an array of bools, of size $|V(G)|$ and mark vertices of a path there. Paths are usually much shorter than $|V(G)|$, but we operate on small graphs, so we can afford this. Second, we notice that we don't need to create this array for each call of \texttt{areUnique} procedure. Let's instead have a static array $stamp$ of ints and a static $counter$ of how many times we called \texttt{areUnique} method. For each element of a path $v_i$, if the value of $stamp$ array is equal to $counter$ we report that vertices are not unique. Else we set $stamp[v_i] = counter$. When returning from the \texttt{areUnique} method we increment $counter$.

@@

The Berge recognition algorithm's running time is $O(n^9)$, which brings into question its applicability to any real use case. Although time complexity is indeed a limiting factor, a number of lower level optimizations done on implementation's level make a very big difference and make it more viable than na\"ive algorithm, at least on some test cases. Also, we explore a new frontier of implementing its most time consuming part on massively parallel GPU architecture, with some good results (\Cref{sec:CUDA}).

@@

Callgrind is a profiling tool that records the call history of a program and presents it as a call-graph. When profiling a program, event counts (data reads, cache misses etc.) are attributed directly to the function that they occurred in. In addition to that, the costs are propagated to functions upwards the call stack. For example, say internal function $worker$ consumes most of programs running time. It is called from $run1$ and $run2$, with $run2$ calls taking twice as much time. $run1$ and $run2$ are in turn called from main. Total attribution of $worker$ would be nearly 100\%, of $run1$ about 33\% and of $run2$ about 66\%. The contribution of main would also be near 100\%. We used a tool \emph{gprof2dot} \cite{gprof2dot} to generate visual call graphs from callgrind's output.

@@

\subsection{Correctness Testing}

@@

In total over 200 cpu-hours of end to end tests were performed without any errors.

@@

CUDA is a parrallel computing platform model created by Nvidia and executed on Nvidia GPUs. It allows to run a program in by many CUDA threads at once. Although usually described by the the PRAM models, implementing PRAM algorithm in CUDA architecture is not as straightforward as implementing a RAM algorith on a cpu. Some of the causes are: synchronization between threads is very costly, there are multiple memory models to choose from, threads run divided in blocks and it is possible to efficiently share memory within blocks and copying memory from and back to CPU suffers from a large latency. What is more, there are no standard dynamic structures in the PRAM model, let alone CUDA, that are taken for granted in the RAM model, such as hashing maps or sets. 
% We used a library called \emph{moderngpu} for our implementation, because it allows a slightly cleaner code.

@@

We posed a question if it would be profitable to use GPU for the problem of recognizing Berge graphs. On one hand the graphs we are working on are small and our time complexity so large that a speedup from massively parallel architecture could be profitable. On the other the algorithm is complex and not easy to parallelise. Therefore we decided to analyze the  CCLSV algorithm and parrallelize parts that would benefit from it the most.

@@

To identify parts of CCLSV to implement, we used callgrind and found potential bottlenecks. The initial tests showed that with growing size of a graph some parts take more and more relative time. Because of callgrind's slow execution times compared to just running the program, we used manual timers for bigger tests. Using this method we identified testing all possible near-cleaners (\Cref{alg:testNearCleaner}) as the biggest bottleneck for larger graphs. We considered two approaches: run whole \Cref{alg:testNearCleaner} on a single CUDA core, testing multiple $X$s in parallel, or parallelize \Cref{alg:testNearCleaner} itself and run it on one $X$ at a time. It turned out that second approach is better and much simpler.

Let us recall the \Cref{alg:testNearCleaner}. It calculates array $R$ of shortest paths, then for each 3-vertex path and an additional vertex it does $O(1)$ checks to see if they along with two paths from $R$ give us an odd hole. We will parallelise the work after calculating $R$, that is, lines \ref{line:cudaStart}-\ref{line:cudaEnd} of the \Cref{alg:testNearCleaner}. Let us notice, that for all $X$s all 3-vertex paths enumerated in line \ref{line:3vertex} are the same. We calculate them beforehand, then each CUDA thread receives one 3-vertex path $x_1 - x_2 - x_3$ and an additional vertex $y_1$ and performs the required checks in lines \ref{line:cudaWorkStard}-\ref{line:cudaWorkEnd}. This is almost perfect scenario for the GPU -- we do a simple SIMD work in parallel, without a lot scattered memory access.

The part of the CCLSV implemented on the gpu achieved speedup of up to 15x. Which gives us a speedup of up to 9x for overall CCLSV algorithm.

>>>>

A repository containing the discussed source code and tests is available under the link: \href{https://github.com/AdrianSiwiec/Perfect-Graphs}{\texttt{https://github.com/AdrianSiwiec/Perfect-Graphs}}
\todo{make repo public and add readme on howto run, some script creating and running default tests}

@@

First of all, there are many methods for generating all possible paths. We could simply enumerate all sequences of vertices without repetition and for each one of them check if it is a path, i.e., if all pairs of vertices next to each other are connected and that there are no other edges, but this would be too slow and render our algorithm unusable.

@@

But we can do much better with some care and a suitable data structure. In addition to having for each vertex a list of its neighbors we create data structure that will allow us to generate a candidate for next path in amortized $O(1)$ time.

@@

In our use case, a few optimizations can be made. First, we don't need a hashing set. We can have an array of bools, of size $|V(G)|$ and mark vertices of a path there. Paths are usually much shorter than $|V(G)|$, but we operate on small graphs, so we can afford this. Second, we notice that we don't need to create this array for each call of \texttt{areUnique} procedure. Let's instead have a static array $stamp$ of ints and a static $counter$ of how many times we called \texttt{areUnique} method. For each element of a path $v_i$, if the value of $stamp$ array is equal to $counter$ we report that vertices are not unique. Else we set $stamp[v_i] = counter$. When returning from \texttt{areUnique} method we increment $counter$.

@@

The Berge recognition algorithm's running time is $O(n^9)$, which brings into question its applicability to any real use case. Although time complexity is indeed a limiting factor, a number of lower level optimizations done on implementation's level make a very big difference and make it more viable than na\"ive algorithm, at least on some test cases. We also explore a new frontier of implementing its most time consuming part on massively parallel GPU architecture, with some good results (\Cref{sec:CUDA}).

@@

% \TODO{Caly ponizszy akapit wymaga przeformulowania/skrocenia, nie trzeba wyjasniac detali dzialania callgrinda - tylko napisac ze zlicza event county wg funkcji + ze mozna to wizualizowac}

Callgrind is a profiling tool that records the call history and event counts ((data reads, cache misses etc.) of a program and presents it as a call-graph. It is then possible to visualise this, we used a tool \emph{gprof2dot} \cite{gprof2dot} to generate visual call graphs from callgrind's output.

@@

\subsection{Correctness testing}

@@

In total over 200 CPU-hours of end to end tests were performed without any errors.

@@

CUDA is a parrallel computing platform model created by Nvidia and executed on Nvidia GPUs. It allows to run a program in by many CUDA threads at once. Although usually described by the the PRAM models, implementing PRAM algorithm in CUDA architecture is not as straightforward as implementing a RAM algorith on a CPU. Some of the causes are: synchronization between threads is very costly, there are multiple memory models to choose from, threads run divided in blocks and it is possible to efficiently share memory within blocks and copying memory from and back to CPU suffers from a large latency. What is more, there are no standard dynamic structures in the PRAM model, let alone CUDA, that are taken for granted in the RAM model, such as hashing maps or sets. 
% We used a library called \emph{moderngpu} for our implementation, because it allows a slightly cleaner code.

@@

We posed a question if it would be profitable to use GPU for the problem of recognizing Berge graphs. On one hand the graphs we are working on are small and our time complexity so large that a speedup from massively parallel architecture could be profitable. On the other hand, the algorithm is complex and not easy to parallelise. Therefore we decided to analyze the  CCLSV algorithm and parrallelize parts that would benefit from it the most.

@@

% \TODO{nizgrabnie, przepisać ponizszy}

We identified testing all possible near-cleaners (\Cref{alg:testNearCleaner}) as the biggest bottleneck for larger graphs. We considered two approaches: run whole \Cref{alg:testNearCleaner} on a single CUDA core, testing multiple $X$s in parallel, or parallelize \Cref{alg:testNearCleaner} itself and run it on one $X$ at a time. It turned out that second approach is better and much simpler.

Let us recall the \Cref{alg:testNearCleaner}. It calculates array $R$ of shortest paths, then for each 3-vertex path and an additional vertex it does $O(1)$ checks to see if they along with two paths from $R$ give us an odd hole. We will parallelise the work after calculating $R$, that is, lines \ref{line:cudaStart}-\ref{line:cudaEnd}. Let us notice, that for all $X$s all 3-vertex paths enumerated in line \ref{line:3vertex} are the same. We calculate them beforehand, then each CUDA thread receives one 3-vertex path $x_1 - x_2 - x_3$ and an additional vertex $y_1$ and performs the required checks in lines \ref{line:cudaWorkStard}-\ref{line:cudaWorkEnd}. This is almost perfect scenario for the GPU -- we do a simple SIMD work in parallel, without a lot scattered memory access.

The part of the CCLSV implemented on the GPU achieved speedup of up to 15x. Which gave us a speedup of up to 9x for overall CCLSV algorithm.

### 2020-09-16 paper/perfectGraphs.tex
many many fixes

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003}, algorithms to solve semi-definite programs grew out of the theory of perfect graphs. We will take a look at semi-definite programs and at perfect graph's relation to Shannon capacity in \Cref{sec:ShannonCapacity}.

\section{Strong Perfect Graph Theorem}
\label{sec:SPGT}

@@

This theorem is a consequence of a stronger result proven by Lovász:
\begin{theorem}[\cite{LL72}, Claim 1 of \cite{GC03}]
	\label{thm:omegaalpha}
	A graph $G$ is perfect if and only if for every induced subgraph $H$, the number of vertices of $H$ is at most $\alpha(H)\omega(H)$.
\end{theorem}
\todo{proof? min. 1 page}
% \begin{proof}
% 	First assume that $G$ is perfect -- for every induced subgraph $H$, $\omega(G) = \chi(H)$. Since the number of vertices of $H$ is at most $\alpha(H)\chi(H)$, the claim follows.

@@

\section{Recognizing Berge Graphs}
\label{sec:recognizingBerge}

@@

\begin{theorem}[part of Theorem 2.2 of \cite{MC05}]
	Procedure \textsc{Contains-Pyramid($G$)} works in $O(|V|^9)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triangles (line \ref{line:pyramidTriangle}) and $O(|V|^3)$ triples $s_1$, $s_2$, $s_3$ (line \ref{line:pyramidTriple}), so lines \ref{line:a}-\ref{line:pyramidEnd} are executed at most $O(|V|^6)$ times.

@@

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \Cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \Cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize}
\end{proof}

@@

\begin{theorem}[part of Theorem 6.4 of \cite{MC05}]
	\Cref{alg:t3} runs in $O(|V|^6)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triples $(v_1, v_2, v_5)$ we enumerate in line \ref{line:t3start}. For each, there are $O(|V|)$ $X$s, so lines \ref{line:t3n4start}-\ref{line:t3n4end} are executed $O(|V|^4)$ times.

@@

Let us notice, that if $X$ is a cleaner for $C$ then $C$ is a clean hole in $G \setminus X$.

@@

\begin{theorem}[Theorem 8.1 of \cite{MC05}]
	\label{thm:amenableHoles}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then every shortest hole in $G$ is amenable.
	% \TODO{Any ideas on what else to say here? List all 9 steps?}
	% \TODO{Dobre pytanie. Te 9 kroków to da się zawrzeć w 1-2 zdaniach każde? Jeśli nie, to tylko napisałbym ogólnikowo, że "jest sekwencja kroków, która to robi, tu jest odnośnik"}
\end{theorem}

@@

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \Cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \Cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \Cref{alg:listNearCleaners} is correct.
\end{proof}

@@

Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The proof of the correctness of the other possible result is more complicated, see section 4 and theorem 5.1 of \cite{MC05}.

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole.
\end{theorem}
\begin{proof}
	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \Cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

@@

% \subsection{Summary}

% \TODO{todo?}

>>>>

Perfect graphs are interesting not only because of their theoretical properties, but they are also used in other areas of study e.g. integrality of polyhedra \cite{Chvtal1975, Chudnovsky2003}, radio channel assignment problem \cite{McDiarmid99, McDiarmid2000} and appear in the theory of Shannon capacity of a graph \cite{Lovasz1979}. Also, as pointed out in \cite{alfonsinPerfect2001, Chudnovsky2003}, algorithms to solve semidefinite programs grew out of the theory of perfect graphs. We will take a look at semidefinite programs and at perfect graph's relation to Shannon capacity in \Cref{sec:ShannonCapacity}.

\section{Strong perfect graph theorem}
\label{sec:SPGT}

@@

This theorem is a consequence of a stronger result proven by Lovász:
\begin{theorem}[\cite{LL72}, Claim 1 of \cite{GC03}]
	\label{thm:omegaalpha}
	A graph $G$ is perfect if and only if for every induced subgraph $H$, the number of vertices of $H$ is at most $\alpha(H)\omega(H)$.
\end{theorem}
% \begin{proof}
% 	First assume that $G$ is perfect -- for every induced subgraph $H$, $\omega(G) = \chi(H)$. Since the number of vertices of $H$ is at most $\alpha(H)\chi(H)$, the claim follows.

@@

\section{Recognizing berge graphs}
\label{sec:recognizingBerge}

@@

\begin{theorem}[part of Theorem 2.2 of \cite{MC05}]
	\label{thm:containsPyramid9}
	Procedure \textsc{Contains-Pyramid($G$)} works in $O(|V|^9)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triangles (line \ref{line:pyramidTriangle}) and $O(|V|^3)$ triples $s_1$, $s_2$, $s_3$ (line \ref{line:pyramidTriple}), so lines \ref{line:a}-\ref{line:pyramidEnd} are executed at most $O(|V|^6)$ times.

@@

\begin{theorem}[Theorem 6.3 of \cite{MC05}]
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item if $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item if $P$ is of length 3, let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \Cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole,
		\item if $P$ is odd with the length of at least 5, it follows from \Cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2-x-p_2-\ldots-p_{n-1}-y-v_2$.
	\end{itemize}
\end{proof}

@@

\begin{theorem}[part of Theorem 6.4 of \cite{MC05}]
	\label{thm:t3Time}
	\Cref{alg:t3} runs in $O(|V|^6)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triples $(v_1, v_2, v_5)$ we enumerate in line \ref{line:t3start}. For each, there are $O(|V|)$ $X$s, so lines \ref{line:t3n4start}-\ref{line:t3n4end} are executed $O(|V|^4)$ times.

@@

Let us notice, that if $X$ is a cleaner for $C$, then $C$ is a clean hole in $G \setminus X$.

@@

\begin{theorem}[Theorem 8.1 of \cite{MC05}]
	\label{thm:amenableHoles}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then every shortest odd hole in $G$ is amenable.
	% \TODO{Any ideas on what else to say here? List all 9 steps?}
	% \TODO{Dobre pytanie. Te 9 kroków to da się zawrzeć w 1-2 zdaniach każde? Jeśli nie, to tylko napisałbym ogólnikowo, że "jest sekwencja kroków, która to robi, tu jest odnośnik"}
\end{theorem}

@@

\begin{theorem}[Theorem 9.2 of \cite{MC05}]
	\label{thm:92}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then the result of the \Cref{alg:listNearCleaners} on the graph $G$ is a set $\mathcal{R}$, such that for every shortest amenable hole $C$, there is a near-cleaner for $C$ in $\mathcal{R}$.
\end{theorem}
\begin{proof}
	Let us suppose that $C$ is an amenable hole in $G$. By \Cref{thm:91}, there is a relevant triple $(a, b, c)$ satisfying that theorem. Let $T$ be the set of all $C$-major vertices not in $X(a,b,c)$. From \Cref{thm:91} we get that $T$ is anticonnected. Since $C$ is amenable, there is an edge $uv$ of $C$ that is $T$-complete, and therefore $T \subseteq N(u, v)$. But then $N(u, v) \cup X(a, b, c) \in \mathcal{R}$ is a near-cleaner for $C$. Therefore the output of the \Cref{alg:listNearCleaners} is correct.
\end{proof}

@@

Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The proof of the correctness of the other possible result is more complicated, see section 4 and Theorem 5.1 of \cite{MC05}.

\begin{theorem}[Theorem 5.1 of \cite{MC05}]
	\label{thm:testNCWorks}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Let $X$ be a subset of $V(G)$. If \Cref{alg:testNearCleaner} reports an odd hole, then $G$ contains an odd hole, else there is no shortest odd hole $C$ in $G$, such that $X$ is a near-cleaner for $C$.
\end{theorem}
\begin{proof}
	Below we will prove that if the \Cref{alg:testNearCleaner} reports an odd hole in $G$, there indeed is one. The rest of the proof is more complicated, see section 4 and Theorem 5.1 of \cite{MC05}.

	Let us suppose that there is a choice of $x_1$, $x_2$, $x_3$, $y_1$ satisfying the three conditions in the \Cref{alg:testNearCleaner} and let $y_2$ and $n$ be defined as in there. We claim that $G$ contains an odd hole.

@@

\subsection{Algorithm for Berge graphs recognition}

To summarize, we present a complete algorithm for recognizing Berge graphs.
\begin{alg}[Determine if a graph is Berge]
	\label{alg:testNearCleaner}
	Input: A graph $G$.

	\noindent Output: Determines if $G$ is Berge.
	\label{alg:bergeness}
\end{alg}

\begin{algorithmic}[1]
	\mProcedure{Check-Berge}{$G$}
		\mIf{\texttt{Contains-Pyramid}$(G)$ \OR \texttt{Contains-Jewel}$(G)$ \OR \lsx 
		\texttt{Contains-T1}$(G)$ \OR \texttt{Contains-T2}$(G)$ \OR \texttt{Contains-T3}$(G)$ \label{line:simpleStart}}
			\ls \RETURN $G$ is not Berge
		\mEndIf
		\mIf{\texttt{Contains-Pyramid}$(\overline{G})$ \OR \texttt{Contains-Jewel}$(\overline{G})$ \OR \lsx 
		\texttt{Contains-T1}$(\overline{G})$ \OR \texttt{Contains-T2}$(\overline{G})$ \OR \texttt{Contains-T3}$(\overline{G})$}
			\ls \RETURN $G$ is not Berge
		\mEndIf \label{line:simpleEnd}
		\ls $\mathcal{R} \gets $\texttt{List-Possible-Near-Cleaners}$(G)$ \label{line:getR}
		\mForEach{$X \in \mathcal{R}$} \label{line:testXStart}
			\mIf{\texttt{Test-Near-Cleaner}$(G, X)$ reports an odd hole}
				\ls \RETURN $G$ is not Berge
			\mEndIf
		\mEndFor \label{line:testXEnd}
		\ls $\mathcal{R'} \gets $\texttt{List-Possible-Near-Cleaners}$(\overline{G})$
		\mForEach{$X \in \mathcal{R'}$}
			\mIf{\texttt{Test-Near-Cleaner}$(\overline{G}, X)$ reports an odd hole}
				\ls \RETURN $G$ is not Berge
			\mEndIf
		\mEndFor
		\ls \RETURN $G$ is Berge
	\mEndProcedure
\end{algorithmic}

\begin{theorem}[Theorem 10.1 of \cite{MC05}]
	Given a graph $G = (V, E)$, \Cref{alg:bergeness} works in $O(|V|^9)$ time and determines is $G$ is Berge.
\end{theorem}
\begin{proof}
	First, in lines \ref{line:simpleStart}-\ref{line:simpleEnd} we test whether $G$ or $\overline{G}$ contain a pyramid, a jewel or a configuration of type $\T_1, \T_2$ or $\T_3$. If it is not the case, then by \Cref{thm:amenableHoles} every shortest odd hole in $G$ and in $\overline{G}$ is amenable.

	By \Cref{thm:92} in line \ref{line:getR} we get a set $\mathcal{R}$, such that if $G$ contains an amenable hole $C$, one of the sets $X \in \mathcal{R}$ is a near-cleaner for $C$. In lines \ref{line:testXStart}-\ref{line:testXEnd} we test each $X \in \mathcal{R}$ and if an odd hole is reported, then by \Cref{thm:testNCWorks} $G$ is not Berge. If for all $X \in \mathcal{R}$ an odd hole is not reported, then by \Cref{thm:amenableHoles}, \Cref{thm:92} and \Cref{thm:testNCWorks} there is no odd hole in $G$. We  run the same procedure on $\overline{G}$ and if again we do not detect an odd hole, then $G$ is Berge and by \Cref{thm:spgt} $G$ is perfect.

	Finally, let us add up the total running time. By \Cref{thm:containsPyramid9} checking if there is a pyramid in $G$ or $\overline{G}$ takes $O(|V|^9)$ time. Testing for jewels takes $O(|V|^7)$ time, testing for a configuration of type $\T_1$ takes $O(|V|^5)$ time and testing for configuration of type $\T_2$ takes $O(|V|^6)$ time. By \Cref{thm:t3Time} testing for a configuration of type $\T_3$ takes $O(|V|^6)$ time. Running \Cref{alg:listNearCleaners} takes $O(|V|^5)$ time and returns $O(|V|^5)$ subsets. For each subset running \Cref{alg:testNearCleaner} takes $O(|V|^4)$ time, so testing all possible near-cleaners takes $O(|V|^9)$ time. Then we repeat on the $\overline{G}$.

	The overall running time is $O(|V|^9)$ with testing for pyramids and testing all possible near-cleaners being algorithms with biggest time complexity.
\end{proof}
